Species persistence under climate and fishing

### 2014-01-31 Writing/MS.tex
small edits

When the climate changes, the area in which organisms can survive and reproduce moves through space. 
This change does not occur in isolation but rather appears on a background of other disturbances. We used an 
integrodifference model explicitly accounting for dispersal, reproduction, a shifting environment, and 
harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
persistence. We found threshold rates of harvesting and of environmental shift such that more extreme harvesting or shift will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior 
of the population.  We then measured the interaction between the stressors and we found low but positive 
synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. 
Finally, we introduced two conservation techniques into simulations of the population model -- threshold 
harvest rules and marine protected areas (MPAs) -- and found that 
these approaches could mitigate the negative interaction of the two stressors.  
\hspace{10cm}

@@

Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008}.  They therefore present an important case study of how disturbances interact in their effects on biological populations.  Further, understanding these interactions will be crucial to managing populations subjected to both of these disturbances.  Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and they are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}.  Species that are likely to undergo or already undergoing shifts in range are also subject to harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.  Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in microcosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}. This empirical work underscores the importance of understanding how range shifts and harvesting interact.

@@

Here we construct an analytical model of a fish population subject to both harvesting and climate change-induced range-shift. We explicitly include two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal. Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.  We find that climate-driven range shifts and fishing interact synergistically, when their effects are measured by absolute drop in biomass.  We also find how the threshold level of harvesting that suffices to drive the population extinct depends on quickly the range is shifting.

@@

\noindent where $\bar{x}\in\left[-\frac{L}{2}, \frac{L}{2}\right]$ describes the position within the patch \citep{ZhouKot2011}. \citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct or not.  To quantify the effects of both a shifting environment and harvesting pressure, we additionally found the total biomass in the equilibrium traveling wave.  For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy

@@

We wanted to understand whether the two stressors interact 
additively, synergistically, or antagonistically \citep{Crainetal2008}. In order to quantify the effect of the 
stressors, we found the total biomass of the population when it reached an equilibrium traveling pulse and 
compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.

>>>>

When the climate changes, the habitat in which organisms can survive and reproduce moves through space. 
This change does not occur in isolation but rather appears on a background of other disturbances. We used an 
integrodifference model explicitly accounting for dispersal, reproduction, to examine how two disturbances, range shift and harvesting, interact and govern population persistence. We found threshold rates of harvesting and of environmental shift such that more increasing either disturbance will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior of the population.  We then measured the interaction between the stressors and we found low but positive synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. Finally, we introduced two conservation techniques into simulations of the population model -- threshold harvest rules and marine protected areas (MPAs) -- and found that these approaches could mitigate, under some circumstances, the negative interaction of the two stressors.  
\hspace{10cm}

@@

Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008}.  They therefore present an important case study of how disturbances interact in their effects on biological populations.  Further, understanding these interactions will be crucial to managing populations subjected to both of these disturbances.  Marine fish are already moving in response to climate change \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and they are projected to continue moving in the future \citep{Kelletal2005, Mackenzieetal2007}.  Species that are likely to undergo or already undergoing shifts in range are also subject to harvesting, in addition to many other disturbances including pollution, ocean acidification, habitat fragmentation, and invasive species \citep{Wilcoveetal1998, Salaetal2000, MEA2005, Pinskyetal2013, Barryetal1995, Nyeetal2009}.  Empirical case studies have identified interactions between overfishing and temperature-driven range shifts that suggest synergy in the magnitude of the disturbance \citep{Lingetal2009} and have demonstrated synergistic effects in microcosm experiments between warming temperatures, harvesting and connectivity \citep{Moraetal2007}. This empirical work underscores the importance of understanding how range shifts and harvesting interact.

@@

Here we construct an analytical model of a fish population subject to both harvesting and climate change-induced range-shift. We explicitly include two mechanistic processes central to species' responses to climate and fishing: reproduction and dispersal. Previous work has highlighted the importance of these two processes and their vulnerability to climate change \citep{Fordhametal2013, Hastingsetal2005}.  We find that climate-driven range shifts and fishing interact synergistically, when their effects are measured by absolute drop in biomass.  We also find how the threshold level of harvesting that suffices to drive the population extinct depends on how quickly the range is shifting.

@@

\noindent where $\bar{x}\in\left[-\frac{L}{2}, \frac{L}{2}\right]$ describes the position within the patch \citep{ZhouKot2011}. \citet{ZhouKot2011} only considered whether a shifting environment will drive a population extinct, however to quantify the effects of both a shifting environment and harvesting pressure we must also find the total biomass in the equilibrium traveling wave.  For a separable kernel, the equilibrium traveling pulse $n^*(x)$ must satisfy

@@

In order to quantify whether the two stressors are interacting additively, synergistically, or antagonistically,  we found the total biomass of the population when it reached an equilibrium traveling pulse and 
compared this equilibrium biomass in the presence and absence of each stressor individually or the two stressors together.
